---
title: 'Class 6: R Functions'
author: "Emma Bernal"
date: "1/23/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Function

## Level 2 heading

### Level 3 heading

```{r}
# this is a silly plot!
plot(1:5)
```

Lets see more about **file import** (i.e. reading files into R). The main read function in base R is `read.table()`

### Table 1
```{r}
t1 <- read.table("test1.txt")
```

We need to add arguments to get this file imported

```{r}
t1 <- read.table("test1.txt", sep=",", header=TRUE)
```

Or I could just use `read.csv` which has the arguments I want in this case!

```{r}
t1 <- read.csv("test1.txt")
t1
```

Let's try the second file.

### Table 2
```{r}
t2 <- read.table("test2.txt")
```

To be able to read the above table properly, we need to add arguments. In this case, the values are separated by `$`.

```{r}
t2 <- read.table("test2.txt", sep="$", header=TRUE)
t2
```

Now, the third file.

### Table 3
```{r}
t3 <- read.table("test3.txt")
t3
```


# Back to Function

Our first example function:

```{r}
add <- function(x, y=1){
  # Sum the input x and y
  x+y
}
```

Let's try using this function

```{r}
add(2, 17)
```

How does this function work with vectors as the input?

```{r}
add(c(1,2,4))
```

It carries out the function for each element in the vector!

We can still use arguments, like y=4. Now it will add **4** to each element.

```{r}
add(c(1,2,4), 4)
```

You can increase the elements of the vector by the elements of another vector.

```{r}
add(c(1,2,4), c(2,3,5))
```

What is this `range()` function?

```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
```

```{r}
range(x)
```

It gives the min and the max!

This is our second function
(*make sure to hit the "play" button to put the function into effect!*):

```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(x)
```

Do some more testing..

```{r}
rescale(1:10)
```


```{r}
# How would we get your function to work in this case? (when there is 'NA' data)
rescale(c(1,2,NA,3,10))
```

Let's use `na.rm=` arguments and set it to `TRUE` to remove the 'NA' data.

```{r}
x <- c(1,2,NA,3,10)
rng <-range(x, na.rm=TRUE)
rng
```

Let's even add it to the function we wrote.

```{r}
rescale_rev <- function(x) {
 rng <-range(x, na.rm=TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale_rev(c(1,2,NA,3,10))
```

Going too far:

```{r}
rescale_extra <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("is it me you are looking for?")
 
 if(plot) {
   print("Don't sing again please!")
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```

```{r}
rescale_extra(x)
```

```{r}
rescale_extra(x, plot=TRUE)
```

### Hands on section B.

Only need to install packages once (`install.packages()` in console)
Then to learn what's in it, use `library()`. 

Package "bio3d" - *Biology Structure Analysis*

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

**Q1.** What type of object is returned from the read.pdb() function? 

```{r}
s1 <- read.pdb("4AKE")
s1
```

```{r}
class(s1)
```

```{r}
str(s1)
```

*A list is returned.*

Amino acids

```{r}
s1$seqres
```

Amino Acid single letter code

```{r}
aa321(s1$seqres)
```



**Q2.** What does the trim.pdb() function do?

```{r}
trim.pdb(s1, chain="A", elety="CA")
```

*It condeses the information.*



**Q3.** What input parameter would turn off the marginal black and grey rectangles in the
plots and what do they represent in this case?

```{r}
plotb3(s1.b, typ="l", ylab="Bfactor")
```

*Removing the `sse=` argument would turn off the black and grey rectangles in the plot.*


**Q4.** What would be a better plot to compare across the different proteins? 

*To combine them onto one plot.*

To do that use `points()`.

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", col="black", ylab="Bfactor")
points(s2.b, typ="l", col="blue")
points(s3.b, typ="l", col="red")
```


**Q5.** Which proteins are more similar to each other in their B-factor trends. How could
you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a
resulting dendrogram plot. Look up the documentation to see what each of these
functions does.


```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

