---
title: "Class 12: Structural Bioinformatics Pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Here we inspect the types of structures in the main database for 3D biomolecular data - the PDB.

> Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

PBD > Analyze > PDB Statisics > Experimental Method and Molecular Types > Download CSV file

```{r}
# read in file

stats <- read.csv("Data Export Summary.csv", row.names=1)

# calculate percentage values for each 'method'

# each row of the column total divided by the sum of the total column

answer <- stats$Total / sum(stats$Total) * 100
names(answer) <- rownames(stats)
round(answer, 2)

```


> Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
#total number of proteins by all different methods

round(sum(stats$Proteins) / sum(stats$Total) * 100, 2)
```

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?


